---
title: "Aviation Traffic Analysis and Monte Carlo Numerical Integration"
author: "Tammar Reza"
date: "2026-04-13"
format: html
execute:
  echo: false
  warning: false
  message: false
---

```{r}
library(tidyverse)
library(scales)

airports <- tribble(
  ~airport, ~year, ~passengers,
  "ATL", 2020, 42918685,
  "ATL", 2021, 75704760,
  "ATL", 2022, 93699630,
  "ATL", 2023, 104653451,
  "ATL", 2024, 108067766,
  "ATL", 2025, 106302208,
  "FRA", 2020, 18768601,
  "FRA", 2021, 24812849,
  "FRA", 2022, 48918482,
  "FRA", 2023, 59355389,
  "FRA", 2024, 61564957,
  "FRA", 2025, 63189666,
  "PKX", 2020, 16091449,
  "PKX", 2021, 25051012,
  "PKX", 2022, 10277623,
  "PKX", 2023, 39410776,
  "PKX", 2024, 49441029,
  "PKX", 2025, 53618949,
  "ORD", 2020, 30860251,
  "ORD", 2021, 54020399,
  "ORD", 2022, 68340619,
  "ORD", 2023, 73894226,
  "ORD", 2024, 80043050,
  "ORD", 2025, 84851825,
  "JFK", 2020, 16630642,
  "JFK", 2021, 30788322,
  "JFK", 2022, 55287693,
  "JFK", 2023, 62464331,
  "JFK", 2024, 63265984,
  "JFK", 2025, 62629455,
  "DEN", 2020, 33741129,
  "DEN", 2021, 58828552,
  "DEN", 2022, 69286461,
  "DEN", 2023, 77837917,
  "DEN", 2024, 82358744,
  "DEN", 2025, 82427962
)

table_data <- airports |>
  pivot_wider(names_from = year, values_from = passengers)
```

---

# Table of Passenger Counts

```{r}
table_data
```

---

# Passenger Traffic Plot

```{r}
library(ggplot2)

ggplot(airports, aes(x = year, y = passengers, color = airport)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = 2020:2025) +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Passenger Traffic (2020–2025)",
    x = "Year",
    y = "Total Passengers",
    color = "Airport"
  ) +
  theme_minimal(base_size = 14)
```

---

Interpretation
---
You can see the sharp rise in numbers due to recovery from low numbers of travellers due to Covid-19 restrictions. You can also see fluctuations for a specific airport year to year and airport to airports in specific years. This shows that airport traffic varies due to many factors.*

# Monte Carlo Data 

```{r}
#| label: monte-carlo-multiples
#| fig-width: 10
#| fig-height: 4

library(tidyverse)

f_mc <- function(x) -(x - 4)^2 + 4

run_sim <- function(n) {
  tibble(
    x = runif(n, 2, 6),
    y = runif(n, 0, 4),
    n_label = paste0("n = ", n)
  ) |>
    mutate(status = ifelse(y <= f_mc(x), "on/below", "above"))
}

mc_results <- bind_rows(
  run_sim(10),
  run_sim(100),
  run_sim(1000),
  run_sim(10000)
)

ggplot() +
  geom_function(fun = f_mc, xlim = c(2, 6), color = "black", linewidth = 1) +
  geom_point(data = mc_results, aes(x, y, color = status), alpha = 0.5, size = 0.7) +
  facet_wrap(
    ~fct_relevel(n_label, "n = 10", "n = 100", "n = 1000", "n = 10000"),
    ncol = 4
  ) +
  scale_color_manual(values = c("above" = "firebrick", "on/below" = "forestgreen")) +
  labs(
    title = "Monte Carlo Integration: Small Multiples Comparison",
    x = "x",
    y = "y"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```
# Analysis

Shows multiple graphs showing how the monte carlo data works with higher n, and how close the number is above or below the true value

# Planning and Prompting GenAI tools

Data Tidying and Visualization Plan
The plan was

1. Create a tribble in R. 
2. Pivot the data wider for a table. 
3. Plot the trends using ggplot2 with comma-formatted labels. 
4.Hide the code in the final output but ensure it is functional."


GenAI Prompting Results
Prompt used (Plan-Based): > "I have a dataset of passenger traffic for six airports (ATL, FRA, PKX, ORD, JFK, DEN) from 2020-2025. Following my plan: 1. Create a tribble in R. 2. Pivot the data wider for a table. 3. Plot the trends using ggplot2 with comma-formatted labels. 4. Hide the code in the final output but ensure it is functional."

Result: The tool (Gemini) provided the R code used in Section 1 of this report. It correctly interpreted the pivot_wider requirement and used scales for the formatting.

Comparison of Plan-Based vs. Generic Prompting
Generic Prompt Result : The generic prompt ("Analyze this airport data in R") produced a basic plot without proper formatting like label_comma or organized "Small Multiples," often leaving the data in a long format that is harder to read in a summary table. It also produced broken code, that took multiple prompts to repair

Comparison Narrative:
The plan-based prompt resulted in a much more "presentation-ready" document. By specifically naming the functions (pivot_wider, label_comma) and the structure (Table vs. Plot), the output required zero manual fixing. In contrast, the generic prompt output usually requires multiple follow-up questions to fix axis labels, handle the data structure, or hide the code chunks.

The primary difference is the output, plan-based approach ensures the AI understands the intended final product, on the other hand the generic prompt doesn’t understand fully, creating a less professional and complete document



# Reflection
Things I learned in this course
1. How to use quarto to format documents with datasets and R code
2. How to use R code with raw datasets to visualize it with graphs



# Code Appendix Aiports

```{r}
#| echo: true
#| warning: false
#| message: false


library(tidyverse)

airports <- tribble(
  ~airport, ~year, ~passengers,
  
  "ATL", 2020, 42918685,
  "ATL", 2021, 75704760,
  "ATL", 2022, 93699630,
  "ATL", 2023, 104653451,
  "ATL", 2024, 108067766,
  "ATL", 2025, 106302208,

  "FRA", 2020, 18768601,
  "FRA", 2021, 24812849,
  "FRA", 2022, 48918482,
  "FRA", 2023, 59355389,
  "FRA", 2024, 61564957,
  "FRA", 2025, 63189666,

  "PKX", 2020, 16091449,
  "PKX", 2021, 25051012,
  "PKX", 2022, 10277623,
  "PKX", 2023, 39410776,
  "PKX", 2024, 49441029,
  "PKX", 2025, 53618949,

  "ORD", 2020, 30860251,
  "ORD", 2021, 54020399,
  "ORD", 2022, 68340619,
  "ORD", 2023, 73894226,
  "ORD", 2024, 80043050,
  "ORD", 2025, 84851825,

  "JFK", 2020, 16630642,
  "JFK", 2021, 30788322,
  "JFK", 2022, 55287693,
  "JFK", 2023, 62464331,
  "JFK", 2024, 63265984,
  "JFK", 2025, 62629455,

  "DEN", 2020, 33741129,
  "DEN", 2021, 58828552,
  "DEN", 2022, 69286461,
  "DEN", 2023, 77837917,
  "DEN", 2024, 82358744,
  "DEN", 2025, 82427962
)

airports

table_data <- airports |>
  pivot_wider(
    names_from = year,
    values_from = passengers
  )

table_data


ggplot(airports, aes(x = year, y = passengers, color = airport)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = 2020:2025) +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Passenger Traffic (2020–2025)",
    x = "Year",
    y = "Total Passengers",
    color = "Airport"
  ) +
  theme_minimal(base_size = 14)

```

# Code Appendix Monte Carlo

```{r}
#| echo: true
#| warning: false
#| message: false


f_mc <- function(x) -(x - 4)^2 + 4

run_sim <- function(n) {
  tibble(
    x = runif(n, 2, 6),
    y = runif(n, 0, 4),
    n_label = paste0("n = ", n)
  ) |> 
    mutate(status = ifelse(y <= f_mc(x), "on/below", "above"))
}

mc_results <- bind_rows(run_sim(10), run_sim(100), run_sim(1000), run_sim(10000))

ggplot() +
  geom_function(fun = f_mc, xlim = c(2, 6), color = "black", linewidth = 1) +
  geom_point(data = mc_results, aes(x, y, color = status), alpha = 0.5, size = 0.7) +
  facet_wrap(~fct_relevel(n_label, "n = 10", "n = 100", "n = 1000"), ncol = 4) +
  scale_color_manual(values = c("above" = "firebrick", "on/below" = "forestgreen")) +
  theme_minimal()

```


# Apendix GenAI Usage
1. Used Google Gemini Free Version
2. Used it 4/15/26
4. Prompt ""I have a dataset of passenger traffic for six airports (ATL, FRA, PKX, ORD, JFK, DEN) from 2020-2025. Following my plan: 1. Create a tribble in R. 2. Pivot the data wider for a table. 3. Plot the trends using ggplot2 with comma-formatted labels. 4. Hide the code in the final output but ensure it is functional." 
Then I pasted the whole section of airport code in for reference
5. Response from AI was:
To help you meet the assignment requirements, I have refined your code based on the plan-based prompt you provided. I noticed that in your previous version, the data was being displayed in its raw format and the plot lacked standard formatting for large numbers.

I have updated the code to ensure the data is tidied into a wide format for the table and that the visualization uses comma-formatted labels for professional readability. I have also added the Monte Carlo simulation using a faceted "small multiples" approach as requested.

Here is the complete .qmd file content. You can copy this entire block into VS Code.